### 2019-06-27 LCAforActiveSonarImaging/document.tex
2019-06-26

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig3_phd.pdf}%
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\\[.5\baselineskip]
   \textbf{Left:} MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.\\[.5\baselineskip]
   \textbf{Right:} MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.}\label{III_mvdr_selected_windows}
\else
\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

@@

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \includegraphics[width=\linewidth]{gfx/fig4.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \\[.5\baselineskip]
   \textbf{Upper bound $\beta$:} Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.
   \\[.5\baselineskip]
   \textbf{Upper bound $\phi$:} As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution.
   }
\else
   \begin{figure*}[t]%
   \includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \newline
   \emph{Upper bound $\beta$:}\hfill
   \parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
   \emph{Upper bound $\phi$:}\hfill
   \parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
   }
\fi%
\label{III_oversampling_mosaic_bounds}
\end{figure*}

@@

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/fig5a.pdf}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/fig5b_phd.pdf}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/buske5a_online.pdf}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/buske5b_online.pdf}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\fi

\ifPhdDoc
   \begin{figure*}[tb]\centering%
   \includegraphics[width=\linewidth]{gfx/fig6.pdf}%
\else
   \begin{figure*}[t]%\centering%
   \includegraphics[width=\textwidth]{gfx/buske6_online.pdf}%
\fi
\caption{\emph{LCA images using different window database sizes}. The window boundaries are the same. The upper left image is identical to a rectangularly weighted DAS. Observe that finer sampling of the $\beta$-range improves sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. However, using more than  $N_\beta=2$ window types and $N_\phi=3$ steering angles makes minimal difference.}\label{III_oversampling_mosaic}
\end{figure*}

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/fig7a.pdf}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/fig7b_phd.pdf}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/buske7a_online.pdf}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/buske7b_online.pdf}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\fi

@@

\ifPhdDoc
\clearpage
\appendix
\renewcommand\thesection{\Roman{section}}
\else
\appendices
\fi

@@

\ifPhdDoc
\typeout{heyheyhey}
   \printbibliography[title=References,heading=subbibliography]
%    \bibliographysty
%    \bibliography{library.bib}
%    \print
\else
\typeout{hohoho}
   \ifBuildBibliography
      \bibliographystyle{IEEEtran}
      \bibliography{references}

>>>>

\ifPhdDoc
   \begin{figure*}[t]\centering%
   \makebox[\linewidth][c]{\includegraphics[width=1.25\linewidth]{gfx/fig3_phd.pdf}}
   \caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\\[.5\baselineskip]
   \textbf{Left:} MVDR sectorscan image of the oiltanker Holmengraa, with 40x40 pixel groups for the shadow, highlight and speckle region of the image indicated with red boxes.\\[.5\baselineskip]
   \textbf{Right:} MVDR window amplitude and phase responses computed from the 40x40 pixel groups. The responses are overlayed each other and the amount of overlap is colored using a logarithmic scale. The size of the pixel groups were chosen ad-hoc for the histograms to be visually invariant to a shift in position within the same region. Since each pixel is pre-delayed into focus the unsteered responses all have their center at broadside. The dashed red lines at -13\,dB and -43\,dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3\,dB in highlight. The phase varies most in the highlight region where we see the highest contrast.}\label{III_mvdr_selected_windows}
\else
\setcounter{topnumber}{1}
\setcounter{dbltopnumber}{1}

@@

\ifPhdDoc
   \begin{figure*}[tbp]\centering%
   \makebox[\linewidth][c]{\includegraphics[width=1.25\linewidth]{gfx/fig4.pdf}}
%    \includegraphics[width=\linewidth]{gfx/fig4.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \\[.5\baselineskip]
   \textbf{Upper bound $\beta$:} Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.
   \\[.5\baselineskip]
   \textbf{Upper bound $\phi$:} As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution.
   }
\else
   \begin{figure*}[t]%
   \includegraphics[width=\textwidth]{gfx/buske4_online.pdf}%
   \caption{\emph{Determining Kaiser parameter boundaries}: LCA images of the resolution test cross and anchor created with a large window database, but where the upper bounds for $\beta$ and $\phi$ are varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two \protect\minus{}3\,dB points of the anchor (red object). These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
   \newline
   \emph{Upper bound $\beta$:}\hfill
   \parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{III_mvdr_selected_windows} suggests that MVDR prefers windows with sidelobe levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
   \emph{Upper bound $\phi$:}\hfill
   \parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
   }
\fi%
\label{III_oversampling_mosaic_bounds}
\end{figure*}

@@

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\makebox[\linewidth][c]{\includegraphics[width=1.25\linewidth]{gfx/fig5a.pdf}}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\makebox[\linewidth][c]{\includegraphics[width=1.25\linewidth]{gfx/fig5b_phd.pdf}}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/buske5a_online.pdf}%%
   \label{III_oversampling_mosaic_bounds_beta}}
   \hfil
   \subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{III_oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/buske5b_online.pdf}%
   \label{III_oversampling_mosaic_bounds_phi}}
   \caption{Kaiser windows chosen for each pixel in the image of the resolution test cross. The underlying image is the one shown in \Fig{III_oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
   \label{III_fig_sim}
   \end{figure*}
\fi

\ifPhdDoc
   \begin{figure*}[tb]\centering%
   \makebox[\linewidth][c]{\includegraphics[width=1.25\linewidth]{gfx/fig6.pdf}}%
\else
   \begin{figure*}[t]%\centering%
   \includegraphics[width=\textwidth]{gfx/buske6_online.pdf}%
\fi
\caption{\emph{LCA images using different window database sizes}. The window boundaries are the same. The upper left image is identical to a rectangularly weighted DAS. Observe that finer sampling of the $\beta$-range improves sidelobe suppression but leaves resolution unchanged. Adding more steering-variations improves both sidelobe suppression and resolution. However, using more than  $N_\beta=2$ window types and $N_\phi=3$ steering angles makes minimal difference.}\label{III_oversampling_mosaic}
\end{figure*}

\ifPhdDoc
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\makebox[\linewidth][c]{\includegraphics[width=1.25\textwidth]{gfx/fig7a.pdf}}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\makebox[\linewidth][c]{\includegraphics[width=1.25\textwidth]{gfx/fig7b_phd.pdf}}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\else
   \begin{figure*}[t]\centering
   \subfloat[\emph{Image quality of approximate point scatterers found in the resolution test cross scene}. Two lateral image cuts are shown in the leftmost figures, along with the response of a rectangularly weighted DAS. Note how the FWHM of LCA sits in between that of DAS and MVDR. Also note how LCA is insensitive to the window type; its performance is near identical whether it uses Kasier or trigonometric windows.]{\includegraphics[width=\textwidth]{gfx/buske7a_online.pdf}%
   \label{III_beamformer_comparison_cross}}
   \hfil
   \subfloat[\emph{Image quality of the full sector Holmengraa scene}. Compared to DAS the adaptive beamformers produce deeper shadows, sharper edges and a higher detail level. The adaptive methods produce nearly identical images.]{\includegraphics[width=\textwidth]{gfx/buske7b_online.pdf}%%
   \label{III_beamformer_comparison_holmengraa}}
   \caption{Comparing image quality of DAS, MVDR and LCA with Kaiser or trigonometric windows.}\label{III_image_quality}
   \end{figure*}
\fi

@@

\ifPhdDoc
\clearpage
\appendix
\renewcommand\thesection{\Alph{section}}
\renewcommand\thesubsection{\Alph{section}.\arabic{subsection}}
\else
\appendices
\fi

@@

\ifPhdDoc
\typeout{heyheyhey}
   \printbibliography[title=References,heading=subbibliography]
   \addcontentsline{toc}{section}{References}
%    \bibliographysty
%    \bibliography{library.bib}
%    \print
\else
\typeout{hohoho}
   \ifBuildBibliography
      \bibliographystyle{IEEEtran}
      \bibliography{references}

### 2019-06-27 ThesisJo/2_publications.tex
2019-06-26

The upcoming papers will investigate various ways in which acoustical imaging systems can be improved by making them smarter in an operational sense. This involves granting them the ability to adapt their behavior to the information at hand, and the processing power needed to handle the information fast enough.

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Intro}\label{sec:intro} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

Common:
\begin{itemize}\setlength\baselineskip{0pt}
\item Adaptive
\item Processing speed
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper I}\label{sec:paperI} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{An Optimized GPU implementation of the MVDR Beamformer for Active Sonar Imaging}\\
J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 40, Issue 2, April 2015

\begin{itemize}\setlength\baselineskip{0pt}
\item Optimization: Minimum variance
\item Solution space: Continuous, closed when made robust
\item Time domain
\item Sonar
\item Computationally complex, $O(N^3)$.
\item Information significance: Minimal temporal sample support (~single snapshot)
\end{itemize}

%This is a popular method that analytically estimates the array's weight set that best comply with the minimum variance criterion. Its analytic nature allows it to chose from an infinite number of possible weight sets, but with a solution space limited by the constraint of unit gain in the look direction. 
%
%The first paper is concerned with applying the Minimum Variance Distortionless Response (MVDR) beamformer to active sonar imaging, with emphasis on performing arithmetic reductions and solving the practical its high computational complexity. It shows that  with emphasis on solving and with the process of implementing this method on a \gls{GPU}. The goal is to improve sonar image quality without suffering the consequences of the MVDR's inherently high computational complexity.
%
%Combined, these steps led to a speedup of an order of magnitude or more compared to its straight-forward \gls{CPU} implementation.
%
%In functionality this method acts as our reference method, but its practicality is limited by a high computational complexity. The paper suggests means for reducing the arithmetic complexity and provide extensive details for fitting the implementation on a \gls{GPU} for a major speedup. It also relates the performance to theoretical maximums and investigates remaining bottlenecks.

%Applying the MVDR-beamformer may not seem like fully adaptive technique 
%The Minimum Variance Distortionless Response (MVDR) beamformer is a technique predominant is a fully adaptive method that 
%
%
%When performing image processing data from phased arrays an inherent question arise 
%In the pursuit of creating images with more detail and less noise from phased-array data, a fully adaptive beamformer such as the the Minimum Variance Distortionless Response (MVDR) can be used in the receive

@@

\begin{itemize}
\item Minimum variance
\item Full search space, continuous
\item Beam-space
\item Ultrasound
\end{itemize}

@@

\begin{itemize}
\item Minimum variance
\item Smaller search space, discrete
\item Element-space
\item sonar
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper IV}\label{sec:paperIV} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{A Real-Time Synthetic Aperture Sonar Simulator for Automatic Target Recognition: Notation and application}\\
J.I. Buskenes, H. Midelfart\\
\textit{IEEE Journal of Oceanic Engineering}\\
Submitted DD.MM.2018%\\\\
%Volume 42, Issue 1, January 2017

\begin{itemize}
\item Template-generation
\item Autonomy
\item Sonar
	\item Incredibly fast template matching possible.
	\item Possibly a powerful tool to pass prior knowledge to deep learning
	\item Extracting the essential information is key 
	\item Exploit hardware resources available
\end{itemize}

\section{Onlook}

Machine cognition:
\begin{itemize}
\item Ethical dilemmas. Decision and actions have consequences. Who's legally responsible for a machine's doing? Itself?
\item Safety. Machine flexibility allow more possible outcomes, and a wider range of security issues to consider.
\end{itemize}

R dim > 10000
- Inverting such matrices better to do with gradient decent

>>>>

%The upcoming papers will investigate various ways in which acoustical imaging systems can be improved by making them smarter in an operational sense. This involves granting them the ability to adapt their behavior to the information at hand, and the processing power needed to handle the information fast enough.

%%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\section{Intro}\label{sec:intro} %
%%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% Classification accuracy
% Logarithmic loss
% Confusion matrix
% Area under curve
% F1 score
% Mean absolute error
% Mean squared error
%
%

We present four journal-papers concerned with the \emph{adaptive} and \emph{hardware accelerated} formation of acoustic images: the first three reconstructs them from acoustic echoes, and the last synthesizes them through acoustic simulation.

%
%Common:
%
%GPUs
%Experimental validation on HISAS1030 data.

%\begin{itemize}\setlength\baselineskip{0pt}
%\item \emph{Optimization method}, or criteria for adaptiveness and constraints;
%\item \emph{Solution space}, or range of possible values for an optimization problem that satisfy a problem's constraints;
%\item \emph{Domain}, or data representation, 
%\item \emph{Application}, 
%\item \emph{Computational complexity}, $O(N^3)$.
%\item Hardware acceleration. 
%\item Information significance: Minimal temporal sample support (~single snapshot)
%\end{itemize}
%
%All:
%- GPU

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper I}\label{sec:paperI} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{An Optimized GPU implementation of the\\ MVDR Beamformer for Active Sonar Imaging}\\
J.I. Buskenes, J.P. Åsen, C.-I.C. Nilsen and A. Austeng\\
\textit{IEEE Journal of Oceanic Engineering}\\
Volume 40, Issue 2, April 2015

This article evaluates time domain MVDR on GPUs for use in active sonar imaging. We refer to MVDR as fully adaptive, because it computes array weights analytically from spatial features of the data, yielding a continuous solution space. Like any data-driven algorithm, it requires statistically distinguishable signal and noise. However, in typical active systems, the signal is highly correlated with the noise and has poor temporal sample support. Therefore, in its unconstrained form, MVDR overfits the data, yields high variance and fails to generate representative images.

We relax MVDR through a combination of temporal and spatial averaging, and regularization. This makes it robust and produces images with slightly more details and less noise than delay-and-sum. The resulting solution space is still continuous, but tighter bounded.

We find that implementing MVDR on GPUs instead of CPUs accelerates it by two orders of magnitude, but requires considerable and delicate efforts to reduce arithmetic operations and fine-tune cache usage. The design is limited to 32 channels due data dependencies that are complicated and not inherently parallel, combined with the relatively limited bandwidth of this hardware architecture.

%
%%\begin{itemize}\setlength\baselineskip{0pt}
%\item Solution space is continuous but strictly constrained due to robustification,
%\item Time domain
%\item Computationally complex, $O(N^3)$
%\item Information significance: Minimal temporal sample support (~single snapshot)
%\end{itemize}

%This is a popular method that analytically estimates the array's weight set that best comply with the minimum variance criterion. Its analytic nature allows it to chose from an infinite number of possible weight sets, but with a solution space limited by the constraint of unit gain in the look direction. 
%

%In functionality this method acts as our reference method, but its practicality is limited by a high computational complexity. The paper suggests means for reducing the arithmetic complexity and provide extensive details for fitting the implementation on a \gls{GPU} for a major speedup. It also relates the performance to theoretical maximums and investigates remaining bottlenecks.

%
%When performing image processing data from phased arrays an inherent question arise 
%In the pursuit of creating images with more detail and less noise from phased-array data, a fully adaptive beamformer such as the the Minimum Variance Distortionless Response (MVDR) can be used in the receive

@@

This article evaluates frequency domain (or beamspace) MVDR on GPUs for use in cardiac ultrasound imaging. These systems typically have real-time requirements and more than 32 channels, so we apply a space reduction technique to reduce the computational effort. Since the transmitted beams are narrow, most acoustic energy falls within three to five center frequencies, and a good approximation follows from just computing those.

We find that beamspace MVDR delivers real-time performance independent of channel-count and an image quality indistinguishable from time domain MVDR.

%The first article implements the fully adaptive MVDR beamformer on a GPU for use in active sonar imaging. Like any data-driven algorithm, it requires data with statistically distinguishable signal and noise. However, in typical active systems, the signal is highly correlated with the noise and has weak temporal sample support. Therefore, MVDR in its unconstrained form overfits the data, yields high variance and fails.
%
%We relax MVDR through a combination of temporal and spatial averaging, and regularization. This makes it robust and produces images of slightly higher quality than delay-and-sum. 
%
%Implementing MVDR on a GPU instead of a CPU led to a speed-up of two orders of magnitude, but required considerable and delicate efforts to reduce arithmetic operations and fine-tune cache usage. The data dependencies are complicated and not inherently parallel.

@@

This article describes LCA, an alternative to MVDR that iterates through precomputed sets of array weights and picks the one that best fulfills the minimum variance criterion. It can be viewed as either a simplification of MVDR or as an adaptive extension to DAS.

We find that constrained MVDR typically generates spatial responses between that of uniform weights and Hamming weights. When LCA is configured with six weight sets covering this range, it matches MVDR in image quality. As a bonus, it is inherently robust and easy to work with, understand and accelerate with GPUs.

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\section{Paper IV}\label{sec:paperIV} %
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\textbf{A Real-Time Synthetic Aperture Sonar Simulator for Automatic Target Recognition: Notation and application}\\
J.I. Buskenes, H. Midelfart\\
\textit{IEEE Journal of Oceanic Engineering}\\
Submitted DD.MM.2019%\\\\
%Volume 42, Issue 1, January 2017

This article describes an acoustic simulator for SAS, implemented on GPUs using OpenGL and OpenCL. Its purpose is to enable AUVs to compare its images on-the-fly to simulations of various known objects, for more responsive, accurate and flexible automatic target recognition (ATR).

We find that simulating objects on-the-fly instead of pre-simulating them consistently improves classification accuracy at false positive rates (FPR) below 20\%, and ties above.

We discuss the design in-depth using a new navigational notation system, extended with topological concepts from robotics and homogeneous coordinates. We find it incredibly helpful for keeping track of complex geometries, numerical representations and higher level abstractions such as manipulating rigid objects.

%Acoustic image reconstruction and synthesis are closely related, and

%
%
%
%\begin{itemize}
%\item Template-generation
%\item Autonomy
%\item Sonar
%	\item Incredibly fast template matching possible.
%	\item Possibly a powerful tool to pass prior knowledge to deep learning
%	\item Extracting the essential information is key 
%	\item Exploit hardware resources available
%\end{itemize}
